### 2012-02-08 docs/architecture/architecture.tex
some fixes from Adam Jacobs

The minimum requirement to
define a new problem would be a {\tt GNUmakefile} which describes how
to build the application and an input file which lists the runtime
parameters.  The problem-specific executable is built in the problem
directory by typing {\tt make}.  Source files are found automatically
by searching the directories listed in the {\tt GNUmakefile}.
Customized versions of any source files placed in the problem-directory
override those with the same name found elsewhere.  Any unique
source files (and not simply a custom version of a file found
elsewhere) needs to be listed in a file call {\tt GPackage.mak}
in the problem-directory.

@@

The runtime parameters for the core \maestro\ algorithm are listed
in {\tt MAESTRO/\_parameters}.  That file is parsed at compile-time
by the {\tt MAESTRO/write\_probin.py} script
(along with any other parameter files listed in the {\tt PROBIN\_PARAMETERS}
variable in the {GNUmakefile}, see \S \ref{sec:makefile}).  The script
outputs the {\tt probin.f90} source file.  Each line
in the {\tt \_parameters} file has the form: 
\vskip 3mm
{\em parameter} \hskip 10em  {\em data-type} \hskip 10em  {\em value} 
\vskip 3mm
\noindent where {\em parameter} is the name of the runtime parameter,
{\em data-type} is one of \{{\tt character}, {\tt real}, {\tt
  integer}, {\tt logical}\}, and the {\em value} specifies the default
value for the runtime parameter.  Comments are indicated by a `{\tt
  \#}' character and a used to produce documentation about the
available runtime parameters.  For the documentation, runtime parameters are grouped together
in the {\tt \_parameters} file into categories.  The category headings
are defined by comments in the {\tt \_parameters} file and any comments
following that heading are placed into that category.  The documentation
is produced by the script {\tt MAESTRO/docs/runtime\_parameters/rp.py}.

@@

Problem-specify runtime parameters should be defined in the
problem-directory in a file called {\tt
  \_parameters}\{{\em.problem-name}\}.  This file should be added
to the {\tt PROBIN\_PARAMETERS} variable in the {\tt GNUmakefile}.

@@

The routines in {\tt MAESTRO/initial\_model/} prepare an initial model
for \maestro.  There are two types of routines here, the first type
take an existing 1D initial model produced somewhere else (e.g.\ a
1D stellar evolution code), and and map it onto a uniform grid, at
the desired resolution, using the equation of state in \maestro, and
using \maestro's discretization of hydrostatic equilibrium.  The second
type generate the initial model internally, by integrating the
condition of hydrostatic equilibrium together with a simplifying
assumption on the energy (e.g. isothermal or isentropic).  In
both cases hydrostatic equilibrium is enforced as:
\begin{equation}
\frac{p_{i+1} - p_i}{\Delta r} = \frac{1}{2} (\rho_i + \rho_{i+1})
g_{i+1/2}
\end{equation}
Here, $g_{i+1/2}$ is the edge-centered gravitational acceleration.

@@

\maestro\ initial models are read in by the {\tt extern/model\_parser}
routines.  This expects the initial model to contain a header giving
the number of variables and their names, followed by rows of data
giving the coordinate and data values at that coordinate.  The initial
model should contain the same species data (as mass fractions) as
defined in the {\tt network} module used by the \maestro\ problem.

>>>>

The minimum requirement to
define a new problem would be a {\tt GNUmakefile} which describes how
to build the application and an input file which lists the runtime
parameters.  The problem-specific executable is built in the problem
directory by typing {\tt make}.  Source files are found automatically
by searching the directories listed in the {\tt GNUmakefile}.
Customized versions of any source files placed in the problem-directory
override those with the same name found elsewhere.  Any unique
source files (and not simply a custom version of a file found
elsewhere) needs to be listed in a file called {\tt GPackage.mak}
in the problem-directory.

@@

The runtime parameters for the core \maestro\ algorithm are listed
in {\tt MAESTRO/\_parameters}.  That file is parsed at compile-time
by the {\tt MAESTRO/write\_probin.py} script
(along with any other parameter files listed in the {\tt PROBIN\_PARAMETERS}
variable in the {GNUmakefile}, see \S \ref{sec:makefile}).  The script
outputs the {\tt probin.f90} source file.  Each line
in the {\tt \_parameters} file has the form: 
\vskip 3mm
{\em parameter} \hskip 10em  {\em data-type} \hskip 10em  {\em value} 
\vskip 3mm
\noindent where {\em parameter} is the name of the runtime parameter,
{\em data-type} is one of \{{\tt character}, {\tt real}, {\tt
  integer}, {\tt logical}\}, and the {\em value} specifies the default
value for the runtime parameter.  Comments are indicated by a `{\tt
  \#}' character and are used to produce documentation about the
available runtime parameters.  For the documentation, runtime parameters are grouped together
in the {\tt \_parameters} file into categories.  The category headings
are defined by comments in the {\tt \_parameters} file and any comments
following that heading are placed into that category.  The documentation
is produced by the script {\tt MAESTRO/docs/runtime\_parameters/rp.py}.

@@

Problem-specific runtime parameters should be defined in the
problem-directory in a file called {\tt
  \_parameters}\{{\em.problem-name}\}.  This file should be added
to the {\tt PROBIN\_PARAMETERS} variable in the {\tt GNUmakefile}.

@@

The routines in {\tt MAESTRO/initial\_model/} prepare an initial model
for \maestro.  There are two types of routines here, the first type
modify an existing 1D initial model produced somewhere else (e.g.\ a
1D stellar evolution code), and map it onto a uniform grid, at
the desired resolution, using the equation of state in \maestro, and
using \maestro's discretization of hydrostatic equilibrium.  The second
type generate the initial model internally, by integrating the
condition of hydrostatic equilibrium together with a simplifying
assumption on the energy (e.g. isothermal or isentropic).  In
both cases hydrostatic equilibrium is enforced as:
\begin{equation}
\frac{p_{i+1} - p_i}{\Delta r} = \frac{1}{2} (\rho_i + \rho_{i+1})
g_{i+1/2}
\end{equation}
Here, $g_{i+1/2}$ is the edge-centered gravitational acceleration.

@@

\maestro\ initial models are read in by the {\tt extern/model\_parser}
routines.  This expects the initial model to contain a header giving
the number of variables and their names, followed by rows of data
giving the coordinate and data values at that coordinate.  The initial
model should contain the same species data (in the form of mass fractions) as
defined in the {\tt network} module used by the \maestro\ problem.

### 2012-02-08 docs/getting_started/getting_started.tex
some fixes from Adam Jacobs

If you shell is {\tt Bash}, add
\begin{verbatim}
export BOXLIB_HOME="/path/to/BoxLib/"
\end{verbatim}
to your {\tt .bashrc}.

@@

\maestro\ is run simply as:
runtime parameters.  The code is run simply as:
\begin{verbatim}
 ./main.Linux.Intel.exe inputs_2d
\end{verbatim}
We can also override the default value of any runtime parameter by specifying
it on the commandline as
\begin{verbatim}
 ./main.Linux.Intel.exe inputs_2d --parameter value
\end{verbatim}

@@

As the code runs, it will output both plotfiles and checkpoints as
well as one of more text diagnostic files ({\tt maestro\_diag.out} by
default) with integral or extrema information (like maximum Mach
number) from each timestep.

@@

Visualization and analysis are done with on the plotfiles.  A
number of in-house and externally developed tools can work 
with \boxlib-formatted plotfiles.

@@

\visit\ is a powerful, DOE-supported visualization tool for 2- and 3D
datasets.  It can do contouring, volume rendering, streamlines, ...\
directly from \boxlib\ plotfiles (to open a plotfile in \visit, point to
the {\tt Header} file inside the plotfile directory).  Details on
\visit\ can be found at:\newline
 {\tt https://wci.llnl.gov/codes/visit/home.html}\,. \newline
The easiest way to get started with \visit\ is to download a precompiled
binary from the \visit\ webpage.

@@

{\tt test2} places 3 hots spots in a plane-parallel atmosphere.
Burning makes these buoyant bubbles which role up.  This problem was
used in \cite{lowMach3} to compare with compressible solvers.

>>>>

If your shell is {\tt Bash}, add
\begin{verbatim}
export BOXLIB_HOME="/path/to/BoxLib/"
\end{verbatim}
to your {\tt .bashrc}.

@@

\maestro\ is run simply as:
\begin{verbatim}
 ./main.Linux.Intel.exe inputs_2d
\end{verbatim}
We can also override the default value of any runtime parameter by specifying
it on the commandline as
\begin{verbatim}
 ./main.Linux.Intel.exe inputs_2d --parameter value
\end{verbatim}

@@

As the code runs, it will output both plotfiles and checkpoints as
well as one or more text diagnostic files ({\tt maestro\_diag.out} by
default) with integral or extrema information (like maximum Mach
number) from each timestep.

@@

Visualization and analysis are done on the plotfiles.  A
number of in-house and externally developed tools can work 
with \boxlib-formatted plotfiles.

@@

\visit\ is a powerful, DOE-supported visualization tool for 2- and 3D
datasets.  It can do contouring, volume rendering, streamlines, ...\, ,
directly from \boxlib\ plotfiles (to open a plotfile in \visit, point to
the {\tt Header} file inside the plotfile directory).  Details on
\visit\ can be found at:\newline
 {\tt https://wci.llnl.gov/codes/visit/home.html}\,. \newline
The easiest way to get started with \visit\ is to download a precompiled
binary from the \visit\ webpage.

@@

{\tt test2} places 3 hots spots in a plane-parallel atmosphere.
Burning makes these bubbles buoyant, and then roll up.  This problem was
used in \cite{lowMach3} to compare with compressible solvers.

### 2012-02-08 docs/introduction/introduction.tex
list of tables
update papers
update referneces
some fixes from Adam Jacobs

\noindent We have several papers that describe applications of the
method to Type Ia supernovae and X-ray bursts:
\begin{itemize}
\item {\em Multidimensional Modeling of Type I X-ray Bursts,}
  C. M. Malone, A. Nonaka, A. S. Almgren, J. B. Bell, \& M. Zingale 2011,
  ApJ, 728, 118 (henceforth ``the XRB paper'')
\item {\em The Convective Phase Preceding Type Ia Supernovae,}
  M. Zingale, A. S. Almgren, J. B. Bell, C. M. Malone, A. Nonaka, \& S. E. Woosley 2011,
  submitted to ApJ (henceforth paper V)
\item {\em High-Resolution Simulations of Convection Preceding Ignition in
  Type Ia Supernovae Using Adaptive Mesh Refinement,}
  A. Nonaka, M. Zingale, A. J. Aspden, A. S. Almgren, J. B. Bell, \& S. E. Woosley
\end{itemize}

@@

Most astrophysical hydrodynamics codes
(e.g.\ CASTRO~\cite{castro} or FLASH~\cite{flash}) solve the
compressible Euler equations, which can be written in the form:
\begin{equation}
U_t + \nabla \cdot F(U) = 0
\end{equation}
where U is the vector of conserved quantities, $U = (\rho, \rho u,
\rho E)$, with $\rho$ the density, $u$ the velocity, and $E$ the total
energy per unit mass.  This system of equations can be expressed 
as a system of advection equations:
\begin{equation}
q_t + A(q) q_x = 0
\end{equation}
where $q$ are called the primitive variables, and $A$ is the
Jacobian, $A \equiv \partial F / \partial U$.  The eigenvalues of the
matrix $A$ are the characteristic speeds---the speeds at which
information propagates.  For the Euler equations, these are $u$ and $u
\pm c$, where $c$ is the sound speed.  Solution methods for the
compressible equations make use of this wave-nature to compute fluxes
at the interfaces of grid cells to update the state in time.  An
excellent introduction to these methods is provided by LeVeque's book
\cite{leveque}.  The timestep for these methods is limited by the time
it takes for maximum characteristic speed to traverse one grid cell.
For very subsonic flows, this means that the timestep is dominated by
the propagation of soundwaves, which may not be important to the
overall dynamics of the flow.

@@

\renewcommand{\arraystretch}{1.5}
%
\begin{center}
\begin{longtable}{|l|p{4.0in}|l|}
\caption[definition of symbols.]{definition of symbols.} \label{table:sym} \\
%
\hline \multicolumn{1}{|c|}{\textbf{symbol}} & 
       \multicolumn{1}{ c|}{\textbf{meaning}} & 
       \multicolumn{1}{ c|}{\textbf{units}} \\ \hline 
\endfirsthead

>>>>

\noindent We have several papers that describe applications of the
method to Type Ia supernovae and X-ray bursts:
\begin{itemize}
\item {\em Multidimensional Modeling of Type I X-ray Bursts,}
  C. M. Malone, A. Nonaka, A. S. Almgren, J. B. Bell, \& M. Zingale 2011,
  ApJ, 728, 118 (henceforth ``the XRB paper'')
\item {\em The Convective Phase Preceding Type Ia Supernovae,}
  M. Zingale, A. S. Almgren, J. B. Bell, C. M. Malone, A. Nonaka, \& S. E. Woosley 2011,
  740, 8 (henceforth paper V)
\item {\em High-Resolution Simulations of Convection Preceding Ignition in
  Type Ia Supernovae Using Adaptive Mesh Refinement,}
  A. Nonaka, M. Zingale, A. J. Aspden, A. S. Almgren, J. B. Bell, \& S. E. Woosley 2012, 
  745, 73
\end{itemize}

@@

Most astrophysical hydrodynamics codes
(e.g.\ CASTRO~\cite{castro} or FLASH~\cite{flash}) solve the
compressible Euler equations, which can be written in the form:
\begin{equation}
U_t + \nabla \cdot F(U) = 0
\end{equation}
where U is the vector of conserved quantities, $U = (\rho, \rho u,
\rho E)$, with $\rho$ the density, $u$ the velocity, and $E$ the total
energy per unit mass.  This system of equations can be expressed 
as a system of advection equations:
\begin{equation}
q_t + A(q) q_x = 0
\end{equation}
where $q$ are called the primitive variables, and $A$ is the
Jacobian, $A \equiv \partial F / \partial U$.  The eigenvalues of the
matrix $A$ are the characteristic speeds---the speeds at which
information propagates.  For the Euler equations, these are $u$ and $u
\pm c$, where $c$ is the sound speed.  Solution methods for the
compressible equations make use of this wave-nature to compute fluxes
at the interfaces of grid cells to update the state in time.  An
excellent introduction to these methods is provided by LeVeque's book
\cite{leveque}.  The timestep for these methods is limited by the time
it takes for the maximum characteristic speed to traverse one grid cell.
For very subsonic flows, this means that the timestep is dominated by
the propagation of soundwaves, which may not be important to the
overall dynamics of the flow.

@@

\renewcommand{\arraystretch}{1.5}
%
\begin{center}
\begin{longtable}{|l|p{4.0in}|l|}
\caption[Definition of symbols.]{Definition of symbols.} \label{table:sym} \\
%
\hline \multicolumn{1}{|c|}{\textbf{symbol}} & 
       \multicolumn{1}{ c|}{\textbf{meaning}} & 
       \multicolumn{1}{ c|}{\textbf{units}} \\ \hline 
\endfirsthead
